Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-15 csf-2017/paper.tex
Spell checking

@@

Finally, the $\store{}{}$ instruction updates the memory to the requested value
if the capability has write authority for the requested location. However, the
instruction is complicated by the presence of \emph{local capabilities}, modeled
after the ones in the CHERI processor~\cite{Watson2015Cheri}. Basically, local
capabilities are special in that they can only be kept in registers, i.e.\ they
cannot be stored to memory. This means that local capabilities can be
\emph{temporarily} given to an adversary, for the duration of an invocation: if
we take care to erase the capability from the register file after control is
passed back to us, they will not have been able to store the capability.
However, there is one exception to the rule above: local capabilities can be
stored to memory for which we have a capability with write-local authority
(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accomodate a stack,
where register contents can be stored, including local capabilities. As long as all
capabilities with write-local authority are themselves local and the stack is
cleared after control is passed back by the adversary, we will see that this
does not break the intended behavior of local capabilities.

We point out that our local capabilities capture only a part of the semantics of
local capabilities in CHERI. Specifically, in addition to the above, CHERI's
default implementation of the CCall exception handler forbids local capabilities
from being passed accross module boundaries. Such a restriction fundamentally
breaks our calling convention, since we pass around local return pointers and
stack capabilities. However, CHERI's CCall is not implemented in hardware, but
in software, precisely to allow experimenting with alternative models like ours.

@@

\textbf{Program startup} We assume that the language runtime initializes the
memory as follows: a contiguous array of memory is reserved for the stack, for
which we receive a stack pointer in a special register $r_\stk$. We stress that
the stack is not built-in, but merely an abstraction we put on this piece of the
memory. The stack pointer is local and has $\rwlx$ permission. Crucially, the
stack is the only part of memory for which the runtime (including malloc,
loading, linking) will ever provide $\rwlx$ or $\rwl$ capabilities.
Additionally, our examples typically also assume some memory to store
instructions or static data. Another part of memory (called the heap) is
initially governed by malloc and at program startup, no other code has
capabilities for this memory. Malloc hands out $\rwx$ capabilities for allocated
regions as requested (no $\rwlx$ or $\rwl$ permissions). For simplicity, we
assume that memory allocated through malloc cannot be freed.

@@

\textbf{Calling convention} With the security measures introduced and motivated,
here is a summary of our proposed calling convention:
\begin{description}[font=\normalfont\itshape]
\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system never hands out global write-local capabilities.
\item[Before returning to the adversary] clear the non-return-value registers and the entire part of the stack that we had access to.
\item[Before invoking the adversary] push an activation record to the stack, create a return pointer as a local $\entry$-capability to the instructions in the activation record, restrict the stack capability to the unused part in $r_\stk$, clear that part of the stack, and clear the non-argument registers.
\item[Before invoking an adversary callback] make sure the callback capability is global.
\item[When called by an adversary] make sure the received stack pointer has permission $\rwlx$.
\end{description}
\textbf{Reusable macro instructions} We define a number of reusable macros
capturing the calling convention and other conveniences. All macros that uses
the stack assumes a stack pointer in register $r_\stk$.
\texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
$\var{name}$ from the linking table and stores it in register $r$.
\texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}} work as one
would expect. \texttt{\footnotesize{prepstk $r$}} is used when a callback is
invoked by the adversary and prepares the received stack pointer by checking
that it has permission $\rwlx$ and making it conform to our calling convention.
\texttt{\footnotesize{scall
    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} jumps to the
capability in register $r$ in the manner described above. That is, it pushes
local state (the contents of registers $\overline{r_{\var{priv}}}$), stack
pointer and return pointer to the stack, sets up the necessary on stack return
code and creates an $\entry$ return pointer, restricts the stack pointer, clears
the unused part of the stack, clears the necessary registers and jumps to $r$.
Upon return, the private state is restored. \texttt{\footnotesize{mclear $r$}}
clears all the memory the capability in register $r$ has authority over.
\texttt{\footnotesize{rclear $\var{regSet}$}} clears all the registers in
$\var{regSet}$. \texttt{\footnotesize{reqglob $r$}} checks whether the word in
register $r$ is a global capability. \texttt{\footnotesize{crtcls
    $\overline{(x_i,r_i)}$ $r$}} allocates a closure where $r$ points to the
closure's code and a new environment is allocated (using malloc) where the
contents of $\overline{r_i}$ is stored. In the code referred to by $r$, an
implicit fetch happens when an instruction refers to $x_i$.

@@

\section{Logical relation}
\label{sec:logical-relation}
In order to reason about programs on the capability machine and prove properties
such as well-bracketedness and encapsulation, we define a unary logical
relation.
% Intuition
It is a semantic model of all the words that are definitely safe to give to an
adversary. Whether a capability is safe depends heavily on the contents of the
memory since instructions as well as capabilities may reside there. Hence we
index the logical relation over worlds, modelling the memory (this makes our
logical relation into a Kripke logical relation). A world consists of a finite
set of regions, and each region defines a protocol that governs the possible
evolution of a part of memory. A future world relation defines how worlds may
evolve during execution (e.g. new protocols established over memory received
from malloc, or steps in a protocol already established for a part of memory).
In fact, as we will explain below, there are two future world relations to
accomodate reasoning about local capabilities.
%

@@

%\dominique{``end up being defined recursively'' does not sound very insightful}
A region defines what its part of memory can legally look like. However,
validity of the memory's contents may itself depend on the state of other parts
of memory. To accomodate this, worlds need to be defined recursively: Worlds
describe memory, which can contain capabilities, whose safety is
world-dependent. Recursive worlds are a common requirement in Kripke models and
the following lemma uses the method of \citet{Birkedal:2011:SKM:1926385.1926401}
for constructing them. The formulation of the lemma is technical, so we
recommend that non-expert readers ignore the technicalities and accept that
there exists a set of worlds $\Wor$ and two relations $\futurestr$ and
$\futurewk$ satisfying the (recursive) equations in the theorem (where the
$\blater$ operator can be safely ignored).
% Short technical explanation
% The recursiveness of the worlds gives rise to a recurisve domain
% equation. Theorem~\ref{thm:world-existence} gives the solution to this
% recursive domain equation.
% \dominique{it is not explained what cofes are?}
% \lau{I added what cofe means but no formal definition (I think this goes under the category technicalities)}
\begin{theorem}\label{thm:world-existence}
  There exists a \cofe{} (complete ordered family of equivalences) $\Wor$ and preorders $\futurestr$ and
  $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are
  preordered \cofes{}, and there exists an isomorphism $\xi$ such that
  \begin{align*}
      \xi &\; : \; \Wor \cong \blater (\nats \finparfun \Regions) \\
      \Regions & = \{\revoked\} \uplus \\
                 &\;\quad \begin{multlined}
                     \{\temp\} \times \States \times \Rels \times \\
                     (\States \fun (\Wor \monwknefun \UPred{\HeapSegments}))
                   \end{multlined} \uplus \\
                 &\;\quad \begin{multlined}
                     \{\perma\} \times \States \times \Rels \times \\
                     (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
                   \end{multlined}
    \end{align*}
  and for $W, W' \in \Wor$
  \begin{align*}
    W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
    W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
  \end{align*}
\end{theorem}
We use the $\Regions$ from Theorem~\ref{thm:world-existence} to define
the worlds:
\[
  \Worlds = \nats \finparfun \Regions
\]

@@

\subsection{Safety of the capability machine}
With the logical relation defined, we can now state the fundamental theorem of
our logical relation: a strong theorem that formalises the guarantees offered by
the capability machine. Essentially, it says that a capability that only grants
safe authority is capability safe as a program counter.
\begin{theorem}[Fundamental Theorem]
  \label{thm:ftlr}
  If one of the following holds:
  \begin{align*}
      & \bullet
        \begin{aligned}[t]
        &\perm = \exec \text{\footnotesize{ and }}\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwx \text{\footnotesize{ and }} \\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwlx \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
      \end{aligned}
  \end{align*}
  then $\npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)$
\end{theorem}
The Fundamental Theorem allows us to prove capability safety for
adveserial programs: if the authority granted by
the capabilty for the untrustred program is reasonable, then it is safe to
execute. Intuitively, this is because our logical relation captures the
safe behaviours of the capability machine that do not allow you to
produce something \emph{unsafe} from something safe. 
The Fundamental Theorem also serves as a sanity check of the
logical relation as it guarantees that a large class of capabilities
actually inhabits it.

@@

To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the start
configuration is capability safe (in the $\observations$ relation) for a world
with a permanent region that requires the assertion flag to be 0. We prove this
using an anti-reduction lemma by which it suffices to show that the
configuration is capability safe after some reduction steps. We then use a general lemma
for reasoning about \texttt{\footnotesize{scall}}, by which it suffices to show that (1) the
configuration that \texttt{\footnotesize{scall}} will jump to is capability safe and (2) that
the configuration just after \texttt{\footnotesize{scall}} is done cleaning up is capability
safe. We use the Fundamental Theorem to reason about the unknown code, but
notice that the Fundamental Theorem says nothing about enter capabilities and
that the adversary capability is an enter capability. Luckily the enter
capability becomes $\exec$ after the jump and then, the Fundamental Theorem
applies.

@@

In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accomodate this in the
proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds for
all the situations where control is passed to the adversary. The potentially
local return pointer received by \texttt{\footnotesize{f4}} from the adversary is safe in
$W_1$, so it can only be used in public future worlds. Let us see how $W_6$ is
constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed
by revoking all temporary regions and adding a $\iota^\pwl$ region for the stack
we pass away and static temporary regions (a region that accepts exactly one
memory) for our local stack as well as the adversary's local stack. The given
worlds $W_3$ and $W_5$ are public future worlds in which the adversary chooses
to return to us. None of these worlds can have masked the temporary regions of
$W_1$ with permanent regions. $W_6$ is constructed by revoking the temporary
regions of $W_5$ and reinstating the temporary regions of $W_1$. This makes
$W_6$ a public future world of $W_1$. It is therefore safe to use the return
pointer to return to the adversary, since we have restored validity of any local
state they might have stashed away.

@@

\emph{Public/private future worlds} A novel aspect of our logical relation is
how we model the temporary nature of local capabilities using public/private
future worlds. The main insight is that the temporary, revokable nature of these
capabilities can be seen as a generalisation of the syntactically-enforced
unstoreable status of evaluation contexts in lambda calculi without control
effects (of which well-bracketed control flow is a consequence). To reason about
code that relies on this (particularly, the original awkward example),
\citet{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
contexts using Kripke worlds with public and private future world relations.
Essentially, they allow relatedness of evaluation contexts to be monotone with
respect to a weaker future world relation (public) than relatedness of values,
formalising the idea that it is safe to make temporary internal state
modifications (private world transititions, that invalidate the continuation,
but not other values) while an expression is performing internal steps, as long
as the code returns to a stable state (i.e. transitions to a public future world
of the original) before returning (i.e. invoking the continuation).\lau{I feel there is an awful lot of parentheses here and that some of them might not be strictly necessary (like explaining temporary internal state modifications and returning = invoking the continuation). Dominique, could you decide whether they are essential to stay true to the contents of the cited paper.}
We
generalise this idea to reason about local capabilities: validity of local
capabilities is allowed to be monotone with respect to a weaker future-world
relation than other values, which we can exploit to distinguish between state
changes that are always safe (public future worlds) and changes that are only
valid if we clear all local capabilities (private future worlds). Our future
world relations are similar to \citet{Dreyer:jfp12}'s (for example, our proof of
the awkward example uses exactly the same state transition system),
but they turn up in an entirely different place in the logical relation: rather
than using public future worlds for the special syntactic category of evaluation
contexts, they are used in the value relation depending on the
locality of the capability at hand. Additionally, our worlds are a bit more
complex because, to allow local memory capabilities and write-local
capabilities, they can contain (revokable) temporary regions that are only
monotonous w.r.t. public future worlds, while DNB's worlds are entirely
permanent.

>>>>

% Uncomment me to disable comments
\renewcommand\lau[1]{}
\renewcommand\dominique[1]{}
\renewcommand\lars[1]{}

@@

Finally, the $\store{}{}$ instruction updates the memory to the requested value
if the capability has write authority for the requested location. However, the
instruction is complicated by the presence of \emph{local capabilities}, modeled
after the ones in the CHERI processor~\cite{Watson2015Cheri}. Basically, local
capabilities are special in that they can only be kept in registers, i.e.\ they
cannot be stored to memory. This means that local capabilities can be
\emph{temporarily} given to an adversary, for the duration of an invocation: if
we take care to erase the capability from the register file after control is
passed back to us, they will not have been able to store the capability.
However, there is one exception to the rule above: local capabilities can be
stored to memory for which we have a capability with write-local authority
(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
where register contents can be stored, including local capabilities. As long as all
capabilities with write-local authority are themselves local and the stack is
cleared after control is passed back by the adversary, we will see that this
does not break the intended behavior of local capabilities.

We point out that our local capabilities capture only a part of the semantics of
local capabilities in CHERI. Specifically, in addition to the above, CHERI's
default implementation of the CCall exception handler forbids local capabilities
from being passed across module boundaries. Such a restriction fundamentally
breaks our calling convention, since we pass around local return pointers and
stack capabilities. However, CHERI's CCall is not implemented in hardware, but
in software, precisely to allow experimenting with alternative models like ours.

@@

\textbf{Program start-up} We assume that the language runtime initializes the
memory as follows: a contiguous array of memory is reserved for the stack, for
which we receive a stack pointer in a special register $r_\stk$. We stress that
the stack is not built-in, but merely an abstraction we put on this piece of the
memory. The stack pointer is local and has $\rwlx$ permission. Crucially, the
stack is the only part of memory for which the runtime (including malloc,
loading, linking) will ever provide $\rwlx$ or $\rwl$ capabilities.
Additionally, our examples typically also assume some memory to store
instructions or static data. Another part of memory (called the heap) is
initially governed by malloc and at program start-up, no other code has
capabilities for this memory. Malloc hands out $\rwx$ capabilities for allocated
regions as requested (no $\rwlx$ or $\rwl$ permissions). For simplicity, we
assume that memory allocated through malloc cannot be freed.

@@

\textbf{Calling convention} With the security measures introduced and motivated,
here is a summary of our proposed calling convention:
\begin{description}[font=\normalfont\itshape]
\item[At program start-up] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system never hands out global write-local capabilities.
\item[Before returning to the adversary] clear the non-return-value registers and the entire part of the stack that we had access to.
\item[Before invoking the adversary] push an activation record to the stack, create a return pointer as a local $\entry$-capability to the instructions in the activation record, restrict the stack capability to the unused part in $r_\stk$, clear that part of the stack, and clear the non-argument registers.
\item[Before invoking an adversary callback] make sure the callback capability is global.
\item[When called by an adversary] make sure the received stack pointer has permission $\rwlx$.
\end{description}
\textbf{Reusable macro instructions} We define a number of reusable macros
capturing the calling convention and other conveniences. All macros that uses
the stack assumes a stack pointer in register $r_\stk$.
\texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
$\var{name}$ from the linking table and stores it in register $r$.
\texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}} work as one
would expect. \texttt{\footnotesize{prepstk $r$}} is used when a callback is
invoked by the adversary and prepares the received stack pointer by checking
that it has permission $\rwlx$ and making it conform to our calling convention.
\texttt{\footnotesize{scall
    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} jumps to the
capability in register $r$ in the manner described above. That is, it pushes
local state (the contents of registers $\overline{r_{\var{priv}}}$), stack
pointer and return pointer to the stack, sets up the necessary on stack return
code and creates an $\entry$ return pointer, restricts the stack pointer, clears
the unused part of the stack, clears the necessary registers and jumps to $r$.
Upon return, the private state is restored. \texttt{\footnotesize{mclear $r$}}
clears all the memory the capability in register $r$ has authority over.
\texttt{\footnotesize{rclear $\var{regSet}$}} clears all the registers in
$\var{regSet}$. \texttt{\footnotesize{reqglob $r$}} checks whether the word in
register $r$ is a global capability. \texttt{\footnotesize{crtcls
    $\overline{(x_i,r_i)}$ $r$}} allocates a closure where $r$ points to the
closure's code and a new environment is allocated (using malloc) where the
contents of $\overline{r_i}$ is stored. In the code referred to by $r$, an
implicit fetch happens when an instruction refers to $x_i$.

@@

\section{Logical relation}
\label{sec:logical-relation}
In order to reason about programs on the capability machine and prove properties
such as well-bracketedness and encapsulation, we define a unary logical
relation.
% Intuition
It is a semantic model of all the words that are definitely safe to give to an
adversary. Whether a capability is safe depends heavily on the contents of the
memory since instructions as well as capabilities may reside there. Hence we
index the logical relation over worlds, modelling the memory (this makes our
logical relation into a Kripke logical relation). A world consists of a finite
set of regions, and each region defines a protocol that governs the possible
evolution of a part of memory. A future world relation defines how worlds may
evolve during execution (e.g. new protocols established over memory received
from malloc, or steps in a protocol already established for a part of memory).
In fact, as we will explain below, there are two future world relations to
accommodate reasoning about local capabilities.
%

@@

%\dominique{``end up being defined recursively'' does not sound very insightful}
A region defines what its part of memory can legally look like. However,
validity of the memory's contents may itself depend on the state of other parts
of memory. To accommodate this, worlds need to be defined recursively: Worlds
describe memory, which can contain capabilities, whose safety is
world-dependent. Recursive worlds are a common requirement in Kripke models and
the following lemma uses the method of \citet{Birkedal:2011:SKM:1926385.1926401}
for constructing them. The formulation of the lemma is technical, so we
recommend that non-expert readers ignore the technicalities and accept that
there exists a set of worlds $\Wor$ and two relations $\futurestr$ and
$\futurewk$ satisfying the (recursive) equations in the theorem (where the
$\blater$ operator can be safely ignored).
% Short technical explanation
% The recursiveness of the worlds gives rise to a recurisve domain
% equation. Theorem~\ref{thm:world-existence} gives the solution to this
% recursive domain equation.
% \dominique{it is not explained what cofes are?}
% \lau{I added what cofe means but no formal definition (I think this goes under the category technicalities)}
\begin{theorem}\label{thm:world-existence}
  There exists a \cofe{} (complete ordered family of equivalences) $\Wor$ and preorders $\futurestr$ and
  $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are
  preordered \cofes{}, and there exists an isomorphism $\xi$ such that
  \begin{align*}
      \xi &\; : \; \Wor \cong \blater (\nats \finparfun \Regions) \\
      \Regions & = \{\revoked\} \uplus \\
                 &\;\quad \begin{multlined}
                     \{\temp\} \times \States \times \Rels \times \\
                     (\States \fun (\Wor \monwknefun \UPred{\HeapSegments}))
                   \end{multlined} \uplus \\
                 &\;\quad \begin{multlined}
                     \{\perma\} \times \States \times \Rels \times \\
                     (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
                   \end{multlined}
    \end{align*}
  and for $W, W' \in \Wor$
  \begin{align*}
    W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
    W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
  \end{align*}
\end{theorem}
We use the $\Regions$ from Theorem~\ref{thm:world-existence} to define
the worlds:
\[
  \Worlds = \nats \finparfun \Regions
\]

@@

\subsection{Safety of the capability machine}
With the logical relation defined, we can now state the fundamental theorem of
our logical relation: a strong theorem that formalises the guarantees offered by
the capability machine. Essentially, it says that a capability that only grants
safe authority is capability safe as a program counter.
\begin{theorem}[Fundamental Theorem]
  \label{thm:ftlr}
  If one of the following holds:
  \begin{align*}
      & \bullet
        \begin{aligned}[t]
        &\perm = \exec \text{\footnotesize{ and }}\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwx \text{\footnotesize{ and }} \\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwlx \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
      \end{aligned}
  \end{align*}
  then $\npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)$
\end{theorem}
The Fundamental Theorem allows us to prove capability safety for
adversarial programs: if the authority granted by
the capability for the adversarial program is reasonable, then it is safe to
execute. Intuitively, this is because our logical relation captures the
safe behaviours of the capability machine that do not allow you to
produce something \emph{unsafe} from something safe. 
The Fundamental Theorem also serves as a sanity check of the
logical relation as it guarantees that a large class of capabilities
actually inhabits it.

@@

To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the start
configuration is capability safe (in the $\observations$ relation) for a world
with a permanent region that requires the assertion flag to be 0. We prove this
using an anti-reduction lemma by which it suffices to show that the
configuration is capability safe after some reduction steps. We then use a general lemma
for reasoning about \texttt{\footnotesize{scall}}, by which it suffices to show that (1) the
configuration that \texttt{\footnotesize{scall}} will jump to is capability safe and (2) that
the configuration just after \texttt{\footnotesize{scall}} is done cleaning up is capability
safe. We use the Fundamental Theorem to reason about the unknown code, but
notice that the Fundamental Theorem says nothing about enter capabilities and
that the adversary capability is an enter capability. Luckily the enter
capability becomes $\exec$ after the jump and then the Fundamental Theorem
applies.

@@

In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in the
proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds for
all the situations where control is passed to the adversary. The potentially
local return pointer received by \texttt{\footnotesize{f4}} from the adversary is safe in
$W_1$, so it can only be used in public future worlds. Let us see how $W_6$ is
constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed
by revoking all temporary regions and adding a $\iota^\pwl$ region for the stack
we pass away and static temporary regions (a region that accepts exactly one
memory) for our local stack as well as the adversary's local stack. The given
worlds $W_3$ and $W_5$ are public future worlds in which the adversary chooses
to return to us. None of these worlds can have masked the temporary regions of
$W_1$ with permanent regions. $W_6$ is constructed by revoking the temporary
regions of $W_5$ and reinstating the temporary regions of $W_1$. This makes
$W_6$ a public future world of $W_1$. It is therefore safe to use the return
pointer to return to the adversary, since we have restored validity of any local
state they might have stashed away.

@@

\emph{Public/private future worlds} A novel aspect of our logical relation is
how we model the temporary nature of local capabilities using public/private
future worlds. The main insight is that the temporary, revokable nature of these
capabilities can be seen as a generalisation of the syntactically-enforced
unstorable status of evaluation contexts in lambda calculi without control
effects (of which well-bracketed control flow is a consequence). To reason about
code that relies on this (particularly, the original awkward example),
\citet{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
contexts using Kripke worlds with public and private future world relations.
Essentially, they allow relatedness of evaluation contexts to be monotone with
respect to a weaker future world relation (public) than relatedness of values,
formalising the idea that it is safe to make temporary internal state
modifications (private world transitions, that invalidate the continuation,
but not other values) while an expression is performing internal steps, as long
as the code returns to a stable state (i.e. transitions to a public future world
of the original) before returning (i.e. invoking the continuation).\lau{I feel there is an awful lot of parentheses here and that some of them might not be strictly necessary (like explaining temporary internal state modifications and returning = invoking the continuation). Dominique, could you decide whether they are essential to stay true to the contents of the cited paper.}
We
generalise this idea to reason about local capabilities: validity of local
capabilities is allowed to be monotone with respect to a weaker future-world
relation than other values, which we can exploit to distinguish between state
changes that are always safe (public future worlds) and changes that are only
valid if we clear all local capabilities (private future worlds). Our future
world relations are similar to \citet{Dreyer:jfp12}'s (for example, our proof of
the awkward example uses exactly the same state transition system),
but they turn up in an entirely different place in the logical relation: rather
than using public future worlds for the special syntactic category of evaluation
contexts, they are used in the value relation depending on the
locality of the capability at hand. Additionally, our worlds are a bit more
complex because, to allow local memory capabilities and write-local
capabilities, they can contain (revokable) temporary regions that are only
monotonous w.r.t. public future worlds, while DNB's worlds are entirely
permanent.
